---
title: "Penguins Size Comparisons"
author: "Danielle Sclafani"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests

Comparing means between two different groups - do we have enough evidence to make us think that the samples were taken from populations with different means
Mass comparison: male & female chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

#Histograms
ggplot(data=chinstraps, aes(x = body_mass_g)) + geom_histogram(bins = 10) +
  facet_wrap(~sex)

##QQ Plot
ggplot(data=chinstraps, aes(sample = body_mass_g)) + geom_qq() +
  facet_wrap(~sex)

## Get descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median (body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )
```

> comparing sizes between male and female chinstrap penguins
Get the vectors of body_mass for female and male chinstrap penguins separately

```{r}
chinstraps_f <- chinstraps %>% 
  filter( sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

### T-tests, a couple different ways

1. is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
```

Are male chinstrap penguions significantly larger in body mass than females? - this is a one sided test, order matters here

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater") # are chinstrap_m greater than chinstrao_females
```

