---
title: "Penguins Size Comparisons"
author: "Danielle Sclafani"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests

Comparing means between two different groups - do we have enough evidence to make us think that the samples were taken from populations with different means
Mass comparison: male & female chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

#Histograms
ggplot(data=chinstraps, aes(x = body_mass_g)) + geom_histogram(bins = 10) +
  facet_wrap(~sex)

##QQ Plot
ggplot(data=chinstraps, aes(sample = body_mass_g)) + geom_qq() +
  facet_wrap(~sex)

## Get descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median (body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )
```

> comparing sizes between male and female chinstrap penguins
Get the vectors of body_mass for female and male chinstrap penguins separately

```{r}
chinstraps_f <- chinstraps %>% 
  filter( sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

### T-tests, a couple different ways

1. is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
```

Are male chinstrap penguions significantly larger in body mass than females? - this is a one sided test, order matters here

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater") # are chinstrap_m greater than chinstrao_females
```

#. Part 2 Beeswarm graph 

Mean flipper lengths between the three species of penguins, females only

### 1. Make a table of counts by species and sex

```{r}
penguin_counts <- penguins %>%
  count(species, sex)

penguin_counts

# never enough to assume n > 30 is a normal sample, always do additional exploration
```


Create a subset of female penguins

```{r}
penguins_f <- penguins %>% 
  filter(sex == "female")

ggplot(data = penguins_f, aes(x = flipper_length_mm)) + geom_histogram(aes(fill = species)) +
  facet_wrap(~species, scales = "free") # each panel has their own x and y axis limits, use when interested in distribution between groups

ggplot(data = penguins_f, aes(sample = flipper_length_mm)) + geom_qq(aes(color = species)) + #change fill aes to color when points
  facet_wrap(~species, scales = "free") # each panel has their own x and y axis limits, use when interested in distribution between groups

```

### Summary Statistics
starting from penguins_f create a summary table that groups by species and returns the following:
- mean of flipper length
- sd of flipper length
- sample size
- variance
- standard error

```{r}
penguins_f_summary <- penguins_f %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper_length = mean(flipper_length_mm),
    sd_flipper_length = sd(flipper_length_mm),
    sample_size = n(),
    se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
    variance_flipper_length = var(flipper_length_mm)
    )

# use caution when using levene test, as sample size gets larger u will get a significant difference, similar to t tests
#levene test for equal variances
# look at Pr(>F) value
leveneTest(flipper_length_mm ~ species, data = penguins_f) # compare variances of flipper lengths for species in the data set penguins_f

```

### Beeswarm plot with mean +/- one sd

```{r}
ggplot() +
  geom_beeswarm(data = penguins_f, 
                aes(x = species, y = flipper_length_mm),
                color = "gray70",
                size = 3,
                alpha = 0.5) +
  geom_point(data = penguins_f_summary, aes(x = species, y = mean_flipper_length), # layering the mean data points on top of the graph 
             size =4) +
  geom_errorbar(data = penguins_f_summary, 
                aes(x = species, 
                    ymax = mean_flipper_length + sd_flipper_length,
                    ymin = mean_flipper_length - sd_flipper_length), # we want error bars to be one sd above the mean, and one sd below the mean
                width = 0.1) + # changes width of error bars
  theme_classic()
```

### One-way ANOVA

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguins_f)
summary(penguin_aov)

# F value is so high, which means variance between groups is large, and within groups are small, there is a very small p-value. here we would reject the null hypothesis, that flipper length for all three species of penguins are the same

# post hoc test- use this to find out which level in the anova is or is not significantly different

TukeyHSD(penguin_aov) # there is a significant difference of flipper length between all three species

```


> for unbalanaced samples -- Use a Type III Anova with the car package:
1. create a linear model

```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguins_f) # creating linear model
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3

# results are pretty darn close, but you would use this in the "wild" when you have unbalanced design, very different sample sizes 
```




